---
title: "IM3 Run Analysis"
author: "Created by: `r Sys.info()['user']`"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document: 
    css: custom.css
    toc: yes
    smart: false
---

```{r setOptions, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# Set up default options for chunks (sections), turn off error or warning messages.
knitr::opts_chunk$set(echo=FALSE, comment=NA, warning=FALSE, message=FALSE, include=TRUE,                                         fig.path=paste0('html_plots/',scenario.name,"-"),dev="png",fig.width=18)
                       


# Load required R libraries
library(data.table)
library(ggplot2)
library(plyr)
library(knitr)
library(rplexos)

# Set figure height and text size options
fig.height.gen.types = 8
fig.height.reserves = 8
fig.height.duration = 8
fig.height.tall= 8
element.text.size = 25

# Create plot theme to use for generic plot template
plot.theme = theme_gray() + 
  theme(axis.text.x = element_text(angle = -90, hjust=0, size=element.text.size),
        axis.title = element_text(size = 20),
        axis.text.y = element_text(size = element.text.size),
        strip.text.x = element_text(size = 25),
        strip.text.y = element_text(size = 25),
        legend.text = element_text(size = 25))
        
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
```




***
## t-tests on Change in Mean Hydro From 1990 Historical Period 

```{r hydro-change-stats}
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
kable(MeanHydro.Test,format.args=list(digits=3))

# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
```






***
## Change in Cost vs. Hydro
```{r change-cost-vs-hydro-periods, fig.width=10}
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

ggplot(data=annual.gen.stats.avg[is.na(Period)==FALSE&Period!=2080&RCP!='hist.'],aes(x=`Increase in Hydro (GWh)`,y=`Increase in Generation Cost ($)`/1e6,color=RCP,size=`Average Hydro (GWh)`,stroke=1.5))+
  geom_point(shape=1)+
  xlab("Deviation in Annual Hydro (GWh)")+
  ylab("Deviation in Annual Generation Cost (M$)")+
  scale_color_manual(breaks = c('rcp45','rcp85'),
                     limits = c('rcp45', 'rcp85'),
                     values=c("blue","red"),
                     labels=c("RCP 4.5 2050","RCP 8.5 2050"))+
  plot.theme+
  scale_size_continuous(label=scales::comma)+
  scale_x_continuous(label=scales::comma, limits=c(-25000,25000))+
  theme(legend.position="right") +
  ylim(c(-1000,1000))+
  geom_text(aes(label=toupper(gsub('_','',GCM)),hjust=1.2,vjust=1.3),
            show.legend=FALSE,,check_overlap=TRUE,size=3)+               
  geom_hline(yintercept=0,linetype='dotdash')+
  geom_vline(xintercept=0,linetype='dotdash')

ggplot(data=annual.gen.stats.avg[is.na(Period)==FALSE&Period!=2080&RCP!='hist.'],aes(x=`Percent Increase in Hydro (%)`,y=`Percent Increase in Cost (%)`,color=RCP,size=`Average Hydro (GWh)`,stroke=1.5))+
  geom_point(shape=1)+
  xlab("Deviation in Annual Hydro (%)")+
  ylab("Deviation in Annual Generation Cost (%)")+
  scale_color_manual(breaks = c('rcp45','rcp85'),
                     limits = c('rcp45', 'rcp85'),
                     values=c("blue","red"),
                     labels=c("RCP 4.5 2050","RCP 8.5 2050"))+
  plot.theme+
  scale_size_continuous(label=scales::comma)+
  scale_x_continuous(label=scales::comma, limits=c(-12,12))+
  theme(legend.position="right") +
  ylim(c(-12,12))+
  geom_text(aes(label=toupper(gsub('_','',GCM)),hjust=1.2,vjust=1.3),
            show.legend=FALSE,,check_overlap=TRUE,size=3)+               
  geom_hline(yintercept=0,linetype='dotdash')+
  geom_vline(xintercept=0,linetype='dotdash')


ggplot(data=annual.gen.stats.avg[is.na(Period)==FALSE&Period!=2080],aes(x=`Increase in Hydro (GWh)`,y=`Increase in Generation Cost ($)`/1e6,color=GCM,size=`Average Hydro (GWh)`,stroke=1.5))+
  geom_point(shape=1)+
  xlab("Deviation in Annual Hydro (GWh)")+
  ylab("Deviation in Annual Generation Cost (M$)")+
  scale_color_manual(breaks = c('ccsm', 'inm_', 'gfdl'),
                     limits = c('ccsm', 'inm_', 'gfdl'),
                     values=c("grey20", "cyan", "magenta"),
                     labels=c("CCSM4","INMCM4","GFDL-CM3"))+
  plot.theme+
  scale_size_continuous(label=scales::comma)+
  scale_x_continuous(label=scales::comma, limits=c(-25000,25000))+
  theme(legend.position="right") +
  ylim(c(-1000,1000))+
  geom_text(aes(label=toupper(gsub('hist.','',gsub('5','.5',RCP))),hjust=1.3,vjust=1),
            show.legend=FALSE,check_overlap=TRUE,size=3)+
  geom_hline(yintercept=0,linetype='dotdash')+
  geom_vline(xintercept=0,linetype='dotdash')

# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
```

***
## Change versus Precip Change
```{r change-response-vs-precip, fig.width=10}
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

ggplot(data=annual.gen.stats.avg[is.na(Period)==FALSE&Period!=2080&RCP!='hist.'],aes(x=`Change in Precip. (%)`,y=`Percent Increase in Hydro (%)`,stroke=1.5,color=GCM),shape=1,size=2)+
  geom_point()+
    geom_smooth(method="lm",se=FALSE,aes(x=`Change in Precip. (%)`,y=`Percent Increase in Hydro (%)`), inherit.aes=FALSE,linetype="dashed",color="black")+
  geom_point(aes(x=`Change in Precip. (%)`,y=-`Percent Increase in Cost (%)`,color=GCM),shape=15,size=2)+   
  geom_smooth(method="lm",se=FALSE,aes(x=`Change in Precip. (%)`,y=-`Percent Increase in Cost (%)`), inherit.aes=FALSE,color="black")+

  
  xlab("Deviation in Precip (%)")+
  ylab("Deviation in Response (%)")+
  scale_color_manual(breaks = c('ccsm', 'inm_', 'gfdl'),
                     limits = c('ccsm', 'inm_', 'gfdl'),
                     values=c("grey", "cyan", "magenta"),
                     labels=c("CCSM4","INMCM4","GFDL-CM3"))+
  plot.theme+theme(legend.position="bottom")+
  scale_x_continuous(label=scales::comma, limits=c(-10,20))+
  theme(legend.position="bottom") +
  ylim(c(-10,20))


# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
```

***
## t-tests on Change in Gen Cost From 1990 Historical Period 
```{r cost-change-stats}
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
kable(GenCost.Test,format.args=list(digits=3))
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
```


***
## Regional Hydro versus Regional Net Gen 

```{r regional-hydro-netgen-two, fig.height=fig.height.tall*2}
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

ggplot(data=net.gen.hydro[TEPPC.Region.gen!="Canada"&TEPPC.Region.gen!="Mexico"&TEPPC.Region.Hydro!="Canada"&TEPPC.Region.Hydro!="Basin"&TEPPC.Region.Hydro!="Rockies"&TEPPC.Region.Hydro!="Southern California",],(aes(x=`Region Hydro`,y=netgen,color=GCM)))+
  geom_point(size=1.5)+
  facet_grid(TEPPC.Region.gen~TEPPC.Region.Hydro,scales="free_x",labeller=as_labeller(GCM_names))+  scale_color_manual(breaks = c('ccsm', 'inm_', 'gfdl'),
                     limits = c('ccsm', 'inm_', 'gfdl'),
                     values=c('#D81B60', '#88B641', '#06BAD5'),
                     labels=c("CCSM4","INMCM4","GFDL-CM3"))+
    
 scale_x_continuous(labels=scales::comma,expand=c(0,0))+
  scale_y_continuous(labels=scales::comma,breaks=c(-100000,100000))+
  plot.theme+xlab('Region Hydro (GWh)')+ylab('Region Net Generation (GWh)')+
  theme(legend.position="bottom")+  expand_limits(x=0)+
        geom_hline(yintercept=0,linetype='dotdash')

# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
```

***
## Region Net Generation versus Hydro Correlation Coefficients
Pearson correlation coefficients are shown and a student's t-test approach is used to determine the significance of the resulting coefficient based on the sample size. Only significant coefficients at 90% confidence are printed (but colors indicate the other values.
```{r netgen-hydro-correlations-regions, fig.height=fig.height.tall*2}
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
ggplot(data=net.gen.hydro.cor.plot,aes(x=`Region Pair`,y=`Region`))+
geom_tile(aes(fill=Correlation))+
  geom_text(data=net.gen.hydro.cor.plot[`P-value`<=0.05],aes(x=`Region Pair`,y=`Region`,label=Correlation))+
  scale_fill_gradient2(low="blue",high="red",limits=c(-1, 1))+
  plot.theme+ggtitle("Pearson Correlation Coefficients, Significant Values Provided")+
  theme(axis.text.x = element_text(angle = -90, hjust=0, size=10),
        axis.text.y = element_text(size = 10))


ggplot(data=net.gen.hydro.cor.plot,aes(x=`Region Pair`,y=`Region`))+
geom_tile(aes(fill=`P-value`))+
  # geom_text(aes(label=`P-value`))+
  scale_fill_gradient2(low="black",mid="green",high="blue",limits=c(-1, 1))+
  plot.theme+ggtitle("p-val Correlation Coefficients")+
  theme(axis.text.x = element_text(angle = -90, hjust=0, size=10),
        axis.text.y = element_text(size = 10))

# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
```

***
## Region Net Generation Correlation Coefficients - Historical
Pearson correlation coefficients are shown and a student's t-test approach is used to determine the significance of the resulting coefficient based on the sample size. Only significant coefficients at 90% confidence are printed (but colors indicate the other values.
```{r netgen-correlations-regions}
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
ggplot(data=net.gen.cor.hist.plot,aes(x=`Region Pair`,y=`Region`))+
geom_tile(aes(fill=Correlation))+
  geom_text(data=net.gen.cor.hist.plot[`P-value`<=0.05],aes(x=`Region Pair`,y=`Region`,label=Correlation))+
  scale_fill_gradient2(low="blue",high="red",limits=c(-1, 1))+
  plot.theme+ggtitle("Pearson Correlation Coefficients, Historical, Significant Values Provided")

ggplot(data=net.gen.cor.hist.plot,aes(x=`Region Pair`,y=`Region`))+
geom_tile(aes(fill=`P-value`))+
  # geom_text(aes(label=`P-value`))+
  scale_fill_gradient2(low="black",mid="green",high="blue",limits=c(-1, 1))+
  plot.theme+ggtitle("p-val Correlation Coefficients, Historical")

# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
```

***
## Region Net Generation Correlation Coefficients - Future
```{r netgen-correlations-regions-future}
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
ggplot(data=net.gen.cor.fut.plot,aes(x=`Region Pair`,y=`Region`))+
geom_tile(aes(fill=Correlation))+
  geom_text(data=net.gen.cor.fut.plot[`P-value`<=0.05],aes(x=`Region Pair`,y=`Region`,label=Correlation))+
  scale_fill_gradient2(low="blue",high="red",limits=c(-1, 1))+
  plot.theme+ggtitle("Pearson Correlation Coefficients, Future, Significant Values Provided")

ggplot(data=net.gen.cor.fut.plot,aes(x=`Region Pair`,y=`Region`))+
geom_tile(aes(fill=`P-value`))+
  # geom_text(aes(label=`P-value`))+
  scale_fill_gradient2(low="black",mid="green",high="blue",limits=c(-1, 1))+
  plot.theme+ggtitle("p-val Correlation Coefficients, Future")

# kable(upper.tri(net.gen.cor.fut,diag=TRUE)*net.gen.cor.fut,format.args=list(big.mark=",",digits=2))
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
```

***
## Gen Cost vs. Hydro and Emulators
```{r gen-vs-hydro, fig.height=fig.height.tall}
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

ggplot(data=annual.gen.stats)+
    geom_point(aes(x=`NW and NCal Hydro`,y=`Cost ($)`/1e9,color=GCM),show.legend = TRUE,size=3)+
    ylab('Generation Cost (B$)') + xlab('Northwest + N CalHydro Generation GWh') +
    scale_color_manual(breaks = c('ccsm', 'inm_', 'gfdl'),
                     limits = c('ccsm', 'inm_', 'gfdl'),
                     values=c('#D81B60', '#88B641', '#06BAD5'),
                     labels=c("CCSM4","INMCM4","GFDL-CM3"))+
   
  scale_x_continuous(label=scales::comma,limits=c(0,250000),expand=c(0,0))+
  scale_y_continuous(limits=c(0,20),expand=c(0,0))+
  plot.theme

summary(emulator_NW_NCal)

# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
```